---
title: "Naive Bayes in R"
author: "Wisnu Adi Nurcahyo"
date: "March 2nd, 2019"

output:
  html_notebook:
    css: presentation.css
---

# Problem Description

Given a data train that contains 160 objects with seven attributes.
The attributes are **age**, **workclass**, **education**, **marital-status**, **occupation**, **relationship**, and **hours-per-week**.
The "output label" is **income** with two categories: `>50K` and `<=50K`.

Build a classification system with Naive Bayes method that's determine the output label for the data test!


# Solution

Here lies my solution with specific built of functions.
The main purpose is not performance and not as minimum lines as possible implementation.
This is for a clear explanation approach.

## Retrieve the Data Train

```{r}
data_train <- read.csv(file = "TrainsetTugas1ML.csv", header = T, sep = ",")
 
data_train
```


## Data Categorization

To make everything easier, we need to categorize our data.
We need to categorize the data within each label.

### Income

First of all, identify all labels from `income` column.

```{r}
levels(data_train$income)
```

There are two labels, `<=50K` and `>50K`.
Simply, we need to rename the label so we could remember it easily.
As for `<=50K`, we will rename it as `low`.
Then as for `>50K`, we will rename it as `high`.

```{r}
bayes.income.high <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$income == ">50K") {
    bayes.income.high <- c(bayes.income.high, index)
  }
}
 
data_train[bayes.income.high, ]
```

```{r}
bayes.income.low <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$income == "<=50K") {
    bayes.income.low <- c(bayes.income.low, index)
  }
}
 
data_train[bayes.income.low, ]
```


### Age

Same as the `income` label above, we identify all labels from `age` column.

```{r}
levels(data_train$age)
```

The labels are quite straightforward so we don't need to rename them.

```{r}
bayes.age.adult <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$age == "adult") {
    bayes.age.adult <- c(bayes.age.adult, index)
  }
}
 
data_train[bayes.age.adult, ]
```

```{r}
bayes.age.old <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$age == "old") {
    bayes.age.old <- c(bayes.age.old, index)
  }
}
 
data_train[bayes.age.old, ]
```

```{r}
bayes.age.young <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$age == "young") {
    bayes.age.young <- c(bayes.age.young, index)
  }
}
 
data_train[bayes.age.young, ]
```


### Workclass

Now, the `workclass` column. Identify all labels that available.

```{r}
levels(data_train$workclass)
```

There are several long labels. We will use the first words only.

```{r}
bayes.workclass.local <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$workclass == "Local-gov") {
    bayes.workclass.local <- c(bayes.workclass.local, index)
  }
}
 
data_train[bayes.workclass.local, ]
```

```{r}
bayes.workclass.private <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$workclass == "Private") {
    bayes.workclass.private <- c(bayes.workclass.private, index)
  }
}

data_train[bayes.workclass.private, ]
```

```{r}
bayes.workclass.self <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$workclass == "Self-emp-not-inc") {
    bayes.workclass.self <- c(bayes.workclass.self, index)
  }
}
 
data_train[bayes.workclass.self, ]
```


### Education

Next is the `education` column. Identify all labels that available.

```{r}
levels(data_train$education)
```

It seems that we need to simplify these labels as well.
Let use `bachelor`, `highschool`, and `college` instead.

```{r}
bayes.education.bachelor <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$education == "Bachelors") {
    bayes.education.bachelor <- c(bayes.education.bachelor, index)
  }
}
 
data_train[bayes.education.bachelor, ]
```

```{r}
bayes.education.highschool <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$education == "HS-grad") {
    bayes.education.highschool <- c(bayes.education.highschool, index)
  }
}
 
data_train[bayes.education.highschool, ]
```

```{r}
bayes.education.college <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$education == "Some-college") {
    bayes.education.college <- c(bayes.education.college, index)
  }
}
 
data_train[bayes.education.college, ]
```


### Marital Status

Same as above, let's see all labels that available on the `marital.status` column.

```{r}
levels(data_train$marital.status)
```

To put it simply, we will use `divorced`, `married`, and `single` instead.

```{r}
bayes.marital.divorced <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$marital.status == "Divorced") {
    bayes.marital.divorced <- c(bayes.marital.divorced, index)
  }
}
 
data_train[bayes.marital.divorced, ]
```

```{r}
bayes.marital.married <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$marital.status == "Married-civ-spouse") {
    bayes.marital.married <- c(bayes.marital.married, index)
  }
}
 
data_train[bayes.marital.married, ]
```

```{r}
bayes.marital.single <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$marital.status == "Never-married") {
    bayes.marital.single <- c(bayes.marital.single, index)
  }
}
 
data_train[bayes.marital.single, ]
```


### Occupation

Another column is `occupation`. Let see how many labels that available.

```{r}
levels(data_train$occupation)
```

We need to simplify the labels.
Use `repair`, `managerial`, and `specialty` instead.

```{r}
bayes.occupation.repair <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$occupation == "Craft-repair") {
    bayes.occupation.repair <- c(bayes.occupation.repair, index)
  }
}
 
data_train[bayes.occupation.repair, ]
```

```{r}
bayes.occupation.managerial <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$occupation == "Exec-managerial") {
    bayes.occupation.managerial <- c(bayes.occupation.managerial, index)
  }
}
 
data_train[bayes.occupation.managerial, ]
```

```{r}
bayes.occupation.specialty <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$occupation == "Prof-specialty") {
    bayes.occupation.specialty <- c(bayes.occupation.specialty, index)
  }
}
 
data_train[bayes.occupation.specialty, ]
```


### Relationship

Now, collects all labels on the `relationship` column.

```{r}
levels(data_train$relationship)
```

Let's simplify these labels.
Use `husband`, `alone`, and `own_child` instead.

```{r}
bayes.relationship.husband <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$relationship == "Husband") {
    bayes.relationship.husband <- c(bayes.relationship.husband, index)
  }
}
 
data_train[bayes.relationship.husband, ]
```

```{r}
bayes.relationship.alone <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$relationship == "Not-in-family") {
    bayes.relationship.alone <- c(bayes.relationship.alone, index)
  }
}
 
data_train[bayes.relationship.alone, ]
```

```{r}
bayes.relationship.own_child <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$relationship == "Own-child") {
    bayes.relationship.own_child <- c(bayes.relationship.own_child, index)
  }
}
 
data_train[bayes.relationship.own_child, ]
```


### Hour per Week

The last one is `hours.per.week` column.

```{r}
levels(data_train$hours.per.week)
```

Since the labels are pretty short and straightforward, we don't need to rename these labels.

```{r}
bayes.hpw.low <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$hours.per.week == "low") {
    bayes.hpw.low <- c(bayes.hpw.low, index)
  }
}
 
data_train[bayes.hpw.low, ]
```

```{r}
bayes.hpw.many <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$hours.per.week == "many") {
    bayes.hpw.many <- c(bayes.hpw.many, index)
  }
}
 
data_train[bayes.hpw.many, ]
```

```{r}
bayes.hpw.normal <- c()
for (index in 1:nrow(data_train)) {
  if (data_train[index, ]$hours.per.week == "normal") {
    bayes.hpw.normal <- c(bayes.hpw.normal, index)
  }
}
 
data_train[bayes.hpw.normal, ]
```


## A Litte Explanation

The categorization above is needed since we will use probability calculation.
Like, how many `husband` that has a `high` income?
We use the result to a multiplication so we need to checks whether an empty set would appears.

```{r}
# Set of age with low income
empty_set <- length(intersect(bayes.income.low, bayes.age.adult)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.age.old)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.age.young)) == 0
 
# Set of age with high income
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.age.adult)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.age.old)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.age.young)) == 0
 
# Set of workclass with low income
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.workclass.local)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.workclass.private)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.workclass.self)) == 0
 
# Set of workclass with high income
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.workclass.local)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.workclass.private)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.workclass.self)) == 0
 
# Set of education with low income
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.education.bachelor)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.education.college)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.education.highschool)) == 0
 
# Set of education with high income
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.education.bachelor)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.education.college)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.education.highschool)) == 0
 
# Set of marital status with low income
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.marital.divorced)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.marital.married)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.marital.single)) == 0
 
# Set of marital status with high income
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.marital.divorced)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.marital.married)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.marital.single)) == 0
 
# Set of occupation with low income
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.occupation.managerial)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.occupation.repair)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.occupation.specialty)) == 0
 
# Set of occupation with high income
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.occupation.managerial)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.occupation.repair)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.occupation.specialty)) == 0
 
# Set of relationship with low income
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.relationship.alone)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.relationship.husband)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.relationship.own_child)) == 0
 
# Set of relationship with high income
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.relationship.alone)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.relationship.husband)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.relationship.own_child)) == 0
 
# Set of hours per week with low income
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.hpw.low)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.hpw.many)) == 0
empty_set <- empty_set | length(intersect(bayes.income.low, bayes.hpw.normal)) == 0
 
# Set of hours per week with high income
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.hpw.low)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.hpw.many)) == 0
empty_set <- empty_set | length(intersect(bayes.income.high, bayes.hpw.normal)) == 0
 
empty_set
```

Luckily, based on the result above, there is no empty set.
Therefore, we don't need the Laplace smoothing at all.


## Bayesian Calculation Functions

After a long way of checks before, we are now ready to create the Bayesian calculation function.

First, build all needed dependencies for calculating certain probability.

```{r}
# A function to calculates the probability of age with certain income
probability_of_age <- function(income, age) {
  data_income <-
    switch (income,
      "high" = bayes.income.high,
      "low" = bayes.income.low
    )
  
  data_age <-
    switch (as.character(age),
      "adult" = bayes.age.adult,
      "old" = bayes.age.old,
      "young" = bayes.age.young
    )
  
  n_data_total <- length(data_income)
  n_data_matched <- length(intersect(data_income, data_age))
  
  return(n_data_matched / n_data_total)
}
 
# A function to calculates the probability of workclass with certain income
probability_of_workclass <- function(income, workclass) {
  data_income <-
    switch (income,
      "high" = bayes.income.high,
      "low" = bayes.income.low
    )
  
  data_workclass <-
    switch (as.character(workclass),
      "Local-gov" = bayes.workclass.local,
      "Private" = bayes.workclass.private,
      "Self-emp-not-inc" = bayes.workclass.self
    )
  
  n_data_total <- length(data_income)
  n_data_matched <- length(intersect(data_income, data_workclass))
  
  return(n_data_matched / n_data_total)
}
 
# A function to calculates the probability of education with certain income
probability_of_education <- function(income, education) {
  data_income <-
    switch (income,
      "high" = bayes.income.high,
      "low" = bayes.income.low
    )
  
  data_education <-
    switch (as.character(education),
      "Bachelors" = bayes.education.bachelor,
      "HS-grad" = bayes.education.highschool,
      "Some-college" = bayes.education.college
    )
  
  n_data_total <- length(data_income)
  n_data_matched <- length(intersect(data_income, data_education))
  
  return(n_data_matched / n_data_total)
}
 
# A function to calculates the probability of marital status with certain income
probability_of_marital <- function(income, marital) {
  data_income <-
    switch (income,
      "high" = bayes.income.high,
      "low" = bayes.income.low
    )
  
  data_marital <-
    switch (as.character(marital),
      "Divorced" = bayes.marital.divorced,
      "Married-civ-spouse" = bayes.marital.married,
      "Never-married" = bayes.marital.single
    )
  
  n_data_total <- length(data_income)
  n_data_matched <- length(intersect(data_income, data_marital))
  
  return(n_data_matched / n_data_total)
}
 
# A function to calculates the probability of occupation with certain income
probability_of_occupation <- function(income, occupation) {
  data_income <-
    switch (income,
      "high" = bayes.income.high,
      "low" = bayes.income.low
    )
  
  data_occupation <-
    switch (as.character(occupation),
      "Craft-repair" = bayes.occupation.repair,
      "Exec-managerial" = bayes.occupation.managerial,
      "Prof-specialty" = bayes.occupation.specialty
    )
  
  n_data_total <- length(data_income)
  n_data_matched <- length(intersect(data_income, data_occupation))
  
  return(n_data_matched / n_data_total)
}
 
# A function to calculates the probability of relationship with certain income
probability_of_relationship <- function(income, relationship) {
  data_income <-
    switch (income,
      "high" = bayes.income.high,
      "low" = bayes.income.low
    )
  
  data_relationship <-
    switch (as.character(relationship),
      "Husband" = bayes.relationship.husband,
      "Not-in-family" = bayes.relationship.alone,
      "Own-child" = bayes.relationship.own_child
    )
  
  n_data_total <- length(data_income)
  n_data_matched <- length(intersect(data_income, data_relationship))
  
  return(n_data_matched / n_data_total)
}
 
# A function to calculates the probability of hours per week with certain income
probability_of_hpw <- function(income, hpw) {
  data_income <-
    switch (income,
      "high" = bayes.income.high,
      "low" = bayes.income.low
    )
  
  data_hpw <-
    switch (as.character(hpw),
      "low" = bayes.hpw.low,
      "many" = bayes.hpw.many,
      "normal" = bayes.hpw.normal
    )
  
  n_data_total <- length(data_income)
  n_data_matched <- length(intersect(data_income, data_hpw))
  
  return(n_data_matched / n_data_total)
}
```

Then, build the bayesian function.

```{r}
predicts_label_with_bayes <- function(data) {
  n_income_low <- length(bayes.income.low)
  n_income_high <- length(bayes.income.high)
  n_income_total <- n_income_low + n_income_high
  
  probability_of_low_income <- 1
  probability_of_low_income <- probability_of_low_income * probability_of_age("low", data$age)
  probability_of_low_income <- probability_of_low_income * probability_of_workclass("low", data$workclass)
  probability_of_low_income <- probability_of_low_income * probability_of_education("low", data$education)
  probability_of_low_income <- probability_of_low_income * probability_of_marital("low", data$marital.status)
  probability_of_low_income <- probability_of_low_income * probability_of_occupation("low", data$occupation)
  probability_of_low_income <- probability_of_low_income * probability_of_relationship("low", data$relationship)
  probability_of_low_income <- probability_of_low_income * probability_of_hpw("low", data$hours.per.week)
  probability_of_low_income <- probability_of_low_income * (n_income_low / n_income_total)
  
  probability_of_high_income <- 1
  probability_of_high_income <- probability_of_high_income * probability_of_age("high", data$age)
  probability_of_high_income <- probability_of_high_income * probability_of_workclass("high", data$workclass)
  probability_of_high_income <- probability_of_high_income * probability_of_education("high", data$education)
  probability_of_high_income <- probability_of_high_income * probability_of_marital("high", data$marital.status)
  probability_of_high_income <- probability_of_high_income * probability_of_occupation("high", data$occupation)
  probability_of_high_income <- probability_of_high_income * probability_of_relationship("high", data$relationship)
  probability_of_high_income <- probability_of_high_income * probability_of_hpw("high", data$hours.per.week)
  probability_of_high_income <- probability_of_high_income * (n_income_high / n_income_total)
  
  result <- if (probability_of_low_income >= probability_of_high_income) {
    "<=50K"
  } else {
    ">50K"
  }
  
  return(result)
}
```

How accurate our machine learning model?

```{r}
matched <- 0
for (index in 1:nrow(data_train)) {
  if (data_train[index, "income"] == predicts_label_with_bayes(data_train[index, ])) {
    matched <- matched + 1
  }
}
 
percentage <- (matched / nrow(data_train)) * 100
paste(as.character(percentage), "%", sep = "")
```


## Execute the Data Test

First, we retrieve the data test.

```{r}
data_test <- read.csv(file = "TestsetTugas1ML.csv")
 
data_test
```

Then, we predicts the `income` label for each row.

```{r}
for (index in 1:nrow(data_test)) {
  predicted_label <- predicts_label_with_bayes(data_test[index, ])
  data_test[index, "income"] <- predicted_label
}
 
data_test
```

Last, stores the result to `TebakanTugas1ML.csv`.

```{r}
write.csv(data_test, file = "TebakanTugas1ML.csv", row.names = F)
```


# Conclusion

The accuration of Bayesian machine learning approach fairly good, we at least reach $80\%$.
This is a common approach of Naive Bayes.
We could improve the accuration with another approach.
Well, machine learning is about modelling and choosing the right approach.

